Investigation of ant cuticle dataset using image texture analysis

### 2022-08-31 paper/main.tex
address comments

\subsection{Deep learning}

@@

@@

Feng et al. \cite{feng_automated_2013} apply an automated system to classify
moth images based on semantic related visual attributes, which are defined as a
pattern on the moth wings. Feng et al. \cite{feng_automated_2013} use a custom
texture descriptor based on the combination of GLCM and \textit{scale-invariant
    feature transform} (SIFT) features \cite{gotlieb_texture_1990,
    lowe_distinctive_2004}. The method is used to classify 50 different moth species
across 8 families \cite{feng_automated_2013}. Their results
\cite{feng_automated_2013} suggest that traditional feature extraction
techniques for the semantic visual attributes of the moth wings are sufficient
for training a classifier to classify an image between 10 randomly selected moth
species. Marques et al. \cite{marques_ant_2018} propose an ensemble-based method
to classify a large number of ant genera. They use ant images sourced from AntWeb
in order to classify 44,806 specimens with at least 1 picture into 57 genera.

@@

In this section, we describe the creation of the custom dataset used in this
research. We use ant head images from AntWeb \cite{perrichot_antweb_2012} and
define two categories for them based on the appearance of the cuticle texture:
\textit{rough} and \textit{smooth}. For each species used, there exists only one
ant head image. We also refer to these ant head images, which describe the pose
of the ant, as ant cuticule images, due to the clear cuticle texture present on
the head. Some randomly selected images from each category are shown in Figures
\ref{fig:smooth-cuticle-texture} and \ref{fig:rough-cuticle-texture}.

@@

Initially, the sculpture identification protocol had 5 subcategories of cuticle
texture: smooth, punctate, striate, reticulate, and tuberous. For simplicity, we
worked only with the two main categories, and the other 4 categories were
combined to the single rough category. The training set identifications were
reviewed together as a group by the assistants. Once training was complete,
assistants were assigned the same genera of ants to identify independently each
week. A weekly meeting was held to discuss identifications and assign new ones.
These identifications were collected in the master spreadsheet and the
identifications were assigned to individual ant species on a majority basis.

@@

\section{K-views, K-means, and deep learning models for experiments}

@@

We evaluate the performance of the models using accuracy, precision, and F1
score. We also apply Grad-CAM with manual inspection to visualize the activation
weights for classified images to show which features lead to the
classification result. Finally, we apply t-SNE to visualize the separation
learned for the model to further analyze the classifications made by the model.

\section{Experimental results and analysis}

>>>>

@@

\subsection{Deep learning for texture analysis}

@@

Feng et al. \cite{feng_automated_2013} apply an automated system to classify
moth images based on semantic related visual attributes, which are defined as a
pattern on the moth wings. Feng et al. \cite{feng_automated_2013} use a custom
texture descriptor based on the combination of GLCM and \textit{scale-invariant
    feature transform} (SIFT) features \cite{gotlieb_texture_1990,
    lowe_distinctive_2004}. The method is used to classify 50 different moth species
across 8 families \cite{feng_automated_2013}. Their results
\cite{feng_automated_2013} suggest that traditional feature extraction
techniques for the semantic visual attributes of the moth wings are sufficient
for training a classifier to classify an image between 10 randomly selected moth
species. Marques et al. \cite{marques_ant_2018} propose an ensemble-based method
to classify a large number of ant genera. They use ant head, profile, and dorsal
images sourced from AntWeb in order to classify 44,806 specimens with at least
one picture into 57 genera. Although the images used in their work are also from
AntWeb, our goal is the classification of the ant cuticle micro sculpturing into
broad groups based solely on the ant head images.

@@

In this section, we describe the creation of the custom dataset used in this
research. We use ant head images from AntWeb \cite{perrichot_antweb_2012} and
define two categories for them based on the appearance of the cuticle texture:
\textit{rough} and \textit{smooth}. The original AntWeb dataset was extended
with these categories for each applicable ant specimen. The extended dataset is
used for classifying the cuticle micro sculpturing into the broad groups. Due to
the large number of ant species, it is quite difficult to obtain lab-quality
specimen photos. For each species used, there exists only one ant head image. We
also refer to these ant head images, which describe the pose of the ant, as ant
cuticule images, due to the clear cuticle texture present on the head. Some
randomly selected images from each category are shown in Figures
\ref{fig:smooth-cuticle-texture} and \ref{fig:rough-cuticle-texture}.

@@

Initially, the sculpture identification protocol had five subcategories of
cuticle texture: smooth, punctate, striate, reticulate, and tuberous. For
simplicity, we worked only with the two main categories, and the other four
categories were combined to the single rough category. The training set
identifications were reviewed together as a group by the assistants. Once
training was complete, assistants were assigned the same genera of ants to
identify independently each week. A weekly meeting was held to discuss
identifications and assign new ones. These identifications were collected in the
master spreadsheet and the identifications were assigned to individual ant
species on a majority basis.

@@

\section{Methods tested}

@@

We evaluate the performance of the models using accuracy, precision, F1 score
and Matthew's correlation coefficient (MCC) \cite{chicco_matthews_2021}. We also
apply Grad-CAM with manual inspection to visualize the activation weights for
classified images to show which features lead to the classification result.
Finally, we apply t-SNE to visualize the separation learned for the model to
further analyze the classifications made by the model.

\section{Results}
